---
title: "Gong et al Treg - Figure 2"
output: html_document
---

## Load packages

```{r}

library(tidyverse)
library(scales)
library(ggrastr)
library(cowplot)
library(lme4)
library(lmerTest)
library(emmeans)
library(ggpubr)
library(nichenetr)
library(Seurat)

Treg_levels <- c("Naive Treg", "Early-activated Treg", "ISG-hi Treg", "TNFR-hi Treg", "Cycling Treg")
APC_levels <- c("Classical Mono", "Non-classical Mono", "C1QC-hi Macro", "HSP-hi Macro", "Hypoxic Macro", "ISG-hi Macro", "NLRP3-hi Macro", "SPP1-hi Macro")

```

## Figure 2A

A total of 15,578 Treg cells were bioinformatically isolated from the PBMC and tumor of HNSCC and MEL scRNA-seq datasets and re-clustered into 5 subclusters.

```{r}

data_F2A <- read.csv("../Data/Gong_et_al_Treg_Fig_2_A.csv")

## Axes ranges

x_min <- min(data_F2A$umap_1)
x_max <- max(data_F2A$umap_1)
y_min <- min(data_F2A$umap_2)
y_max <- max(data_F2A$umap_2)

## Define the order of cell types

data_F2A$Treg_subclusters <- factor(data_F2A$Treg_subclusters, levels = Treg_levels)

## Define the color panel for each cell type: hue_pal() is the palette for Seurat

colors <- setNames(hue_pal()(length(levels(data_F2A$Treg_subclusters))), 
                   levels(data_F2A$Treg_subclusters))

## Split UMAP: generate each panel without legend

data_F2A_PBMC <- data_F2A %>% 
  filter(sample_type == "PBMC")
data_F2A_TIL <- data_F2A %>% 
  filter(sample_type == "TIL")

UMAP_PBMC <- ggplot(data_F2A_PBMC, aes(x = umap_1, y = umap_2, color = Treg_subclusters)) + 
  rasterize(geom_point(colour = "black", size = 1.2), dpi = 300) +
  rasterize(geom_point(size = 0.6, alpha = 0.8), dpi = 300) + 
  scale_color_manual(values = colors, 
                     name = NULL) + 
  coord_fixed(ratio = 1) + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        text = element_text(), 
        axis.title = element_blank(), 
        axis.text = element_blank(), 
        legend.position = "none") + 
  annotate("segment", x = x_min, xend = x_min + 1.5, y = y_min, yend = y_min, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") +
  annotate("segment", x = x_min, xend = x_min, y = y_min, yend = y_min + 1.5, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") + 
  annotate("text", x = x_min + 1.7, y = y_min, 
           label = "UMAP_1", hjust = 0, vjust = 0.5, size = 4) +
  annotate("text", x = x_min, y = y_min + 1.7, 
           label = "UMAP_2", angle = 90, hjust = 0, vjust = 0.5, size = 4) + 
  annotate("text", x = x_min, y = y_max, 
           label = "PBMC", hjust = 0, vjust = 0.5, size = 4)

UMAP_TIL <- ggplot(data_F2A_TIL, aes(x = umap_1, y = umap_2, color = Treg_subclusters)) + 
  rasterize(geom_point(colour = "black", size = 1.2), dpi = 300) +
  rasterize(geom_point(size = 0.6, alpha = 0.8), dpi = 300) + 
  scale_color_manual(values = colors, 
                     name = NULL) + 
  coord_fixed(ratio = 1) + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        text = element_text(), 
        axis.title = element_blank(), 
        axis.text = element_blank(), 
        legend.position = "none") + 
  annotate("segment", x = x_min, xend = x_min + 1.5, y = y_min, yend = y_min, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") +
  annotate("segment", x = x_min, xend = x_min, y = y_min, yend = y_min + 1.5, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") + 
  annotate("text", x = x_min + 1.7, y = y_min, 
           label = "UMAP_1", hjust = 0, vjust = 0.5, size = 4) +
  annotate("text", x = x_min, y = y_min + 1.7, 
           label = "UMAP_2", angle = 90, hjust = 0, vjust = 0.5, size = 4) + 
  annotate("text", x = x_min, y = y_max, 
           label = "Tumor", hjust = 0, vjust = 0.5, size = 4)

## Extract legend from main UMAP: to ensure all subclusters are included

UMAP <- ggplot(data_F2A, aes(x = umap_1, y = umap_2, color = Treg_subclusters)) + 
  rasterize(geom_point(color = "black", size = 1.2), dpi = 300) +
  rasterize(geom_point(size = 0.6, alpha = 0.8), dpi = 300) + 
  scale_color_manual(values = colors, 
                     name = NULL) + 
  coord_fixed(ratio = 1) + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        text = element_text(color = "black"), 
        axis.title = element_blank(), 
        axis.text = element_blank(), 
        legend.title = element_text(size = 10),
        legend.text  = element_text(size = 8)) + 
  guides(colour = guide_legend(title = "Treg subclusters", override.aes = list(size = 4, alpha = 1)))

UMAP_legend <- cowplot::get_legend(UMAP)

F2A <- plot_grid(plot_grid(UMAP_PBMC, UMAP_TIL, nrow = 1, align = "hv"),
                 UMAP_legend, ncol = 2, rel_widths = c(1, 0.25))
F2A

```

## Figure 2B

UMAP of Treg cells embedded with the mean pseudotemporal value of each cell.

```{r}

data_F2B <- read.csv("../Data/Gong_et_al_Treg_Fig_2_B.csv")

x_min <- min(data_F2B$umap_1)
x_max <- max(data_F2B$umap_1)
y_min <- min(data_F2B$umap_2)
y_max <- max(data_F2B$umap_2)

F2B <- ggplot(data_F2B, aes(x = umap_1, y = umap_2, color = Avg_pseudotime_value)) + 
  rasterize(geom_point(size = 0.6), dpi = 300) + 
  scale_color_gradientn(colors = scales::alpha(viridis::magma(20)[5:20], 0.9), 
                        name = "Pseudotime", 
                        limits = c(0, 100), 
                        breaks = c(0, 25, 50, 75, 100), 
                        oob = scales::squish) + 
  coord_fixed(ratio = 1) + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        axis.title = element_blank(), 
        axis.text = element_blank()) + 
  annotate("segment", x = x_min, xend = x_min + 1.5, y = y_min, yend = y_min, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") +
  annotate("segment", x = x_min, xend = x_min, y = y_min, yend = y_min + 1.5, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") + 
  annotate("text", x = x_min + 2, y = y_min, 
           label = "UMAP_1", hjust = 0, vjust = 0.5, size = 4) +
  annotate("text", x = x_min, y = y_min + 2, 
           label = "UMAP_2", angle = 90, hjust = 0, vjust = 0.5, size = 4)
F2B

```

## Figure 2C

Boxplot comparing the average expression level of CSF1 between TNFRhi Treg and other Treg subsets in each patient.

```{r}

data_F2C <- read.csv("../Data/Gong_et_al_Treg_Fig_2_C.csv")

data_F2C$Treg_subclusters <- factor(
  data_F2C$Treg_subclusters,
  levels = Treg_levels)

gene <- "CSF1"

comparisons <- list(
  c("Naive Treg", "TNFR-hi Treg"),
  c("Early-activated Treg", "TNFR-hi Treg"),
  c("ISG-hi Treg", "TNFR-hi Treg"),
  c("Cycling Treg", "TNFR-hi Treg")
)

## Linear mixed model

model <- lmer(expression ~ Treg_subclusters + (1 | patient_id), data = data_F2C)
anova(model)
emm <- emmeans(model, ~ Treg_subclusters)

## Tukey's multiple comparison

pairwise_comparison <- pairs(emm, adjust = "tukey")
pairwise_result <- as.data.frame(pairwise_comparison) %>%
  separate(contrast, into = c("group1", "group2"), sep = " - ") %>% 
  mutate(group1 = str_remove_all(group1, "[()]"), 
         group2 = str_remove_all(group2, "[()]")) %>% 
  filter(str_detect(group1, "TNFR-hi Treg") |
         str_detect(group2, "TNFR-hi Treg")) %>% 
  mutate(p.label = sprintf("%.2e", p.value))

## Box plot

F2C <- ggplot(data_F2C, aes(x = Treg_subclusters, y = expression, fill = Treg_subclusters)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(aes(fill = Treg_subclusters),
              shape = 21,
              color = "black",
              width = 0.3,
              size = 1,
              alpha = 0.9) + 
  scale_fill_manual(values = hue_pal()(5)) + 
  theme_classic() +
  xlab("") + 
  ylab(paste0("Normalized ", gene, " expression Level")) + 
  theme(axis.text.x = element_text(color = "black", angle = 0, hjust = 0.5), 
        axis.text.y = element_text(color = "black"), 
        axis.line = element_line(color = "black"), 
        axis.ticks = element_line(color = "black"), 
        legend.position = "none") + 
  stat_pvalue_manual(pairwise_result,
                     label = "p.label",
                     xmin = "group1",
                     xmax = "group2",
                     y.position = c(1.3, 1.5, 1.7, 1.9), 
                     inherit.aes = F)
F2C

```

## Figure 2D

A total of 44,296 monocytes and macrophages were bioinformatically isolated from the PBMC and tumor of HNSCC and MEL scRNA-seq datasets and re-clustered into 8 subclusters.

```{r}

data_F2D <- read.csv("../Data/Gong_et_al_Treg_Fig_2_D.csv")

## Axes ranges

x_min <- min(data_F2D$umap_1)
x_max <- max(data_F2D$umap_1)
y_min <- min(data_F2D$umap_2)
y_max <- max(data_F2D$umap_2)

## Define the order of cell types

data_F2D$APC_subclusters <- factor(data_F2D$APC_subclusters, levels = APC_levels)

## Define the color panel for each cell type: hue_pal() is the palette for Seurat

colors <- setNames(hue_pal()(length(levels(data_F2D$APC_subclusters))), 
                   levels(data_F2D$APC_subclusters))

## Split UMAP: generate each panel without legend

data_F2D_PBMC <- data_F2D %>% 
  filter(sample_type == "PBMC")
data_F2D_TIL <- data_F2D %>% 
  filter(sample_type == "TIL")

UMAP_PBMC <- ggplot(data_F2D_PBMC, aes(x = umap_1, y = umap_2, color = APC_subclusters)) + 
  rasterize(geom_point(colour = "black", size = 1.2), dpi = 300) +
  rasterize(geom_point(size = 0.6, alpha = 0.8), dpi = 300) + 
  scale_color_manual(values = colors, 
                     name = NULL) + 
  coord_fixed(ratio = 1) + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        text = element_text(), 
        axis.title = element_blank(), 
        axis.text = element_blank(), 
        legend.position = "none") + 
  annotate("segment", x = x_min, xend = x_min + 1.5, y = y_min, yend = y_min, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") +
  annotate("segment", x = x_min, xend = x_min, y = y_min, yend = y_min + 1.5, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") + 
  annotate("text", x = x_min + 1.7, y = y_min, 
           label = "UMAP_1", hjust = 0, vjust = 0.5, size = 4) +
  annotate("text", x = x_min, y = y_min + 1.7, 
           label = "UMAP_2", angle = 90, hjust = 0, vjust = 0.5, size = 4) + 
  annotate("text", x = x_min, y = y_max, 
           label = "PBMC", hjust = 0, vjust = 0.5, size = 4)

UMAP_TIL <- ggplot(data_F2D_TIL, aes(x = umap_1, y = umap_2, color = APC_subclusters)) + 
  rasterize(geom_point(colour = "black", size = 1.2), dpi = 300) +
  rasterize(geom_point(size = 0.6, alpha = 0.8), dpi = 300) + 
  scale_color_manual(values = colors, 
                     name = NULL) + 
  coord_fixed(ratio = 1) + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        text = element_text(), 
        axis.title = element_blank(), 
        axis.text = element_blank(), 
        legend.position = "none") + 
  annotate("segment", x = x_min, xend = x_min + 1.5, y = y_min, yend = y_min, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") +
  annotate("segment", x = x_min, xend = x_min, y = y_min, yend = y_min + 1.5, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") + 
  annotate("text", x = x_min + 1.7, y = y_min, 
           label = "UMAP_1", hjust = 0, vjust = 0.5, size = 4) +
  annotate("text", x = x_min, y = y_min + 1.7, 
           label = "UMAP_2", angle = 90, hjust = 0, vjust = 0.5, size = 4) + 
  annotate("text", x = x_min, y = y_max, 
           label = "Tumor", hjust = 0, vjust = 0.5, size = 4)

## Extract legend from main UMAP: to ensure all subclusters are included

UMAP <- ggplot(data_F2D, aes(x = umap_1, y = umap_2, color = APC_subclusters)) + 
  rasterize(geom_point(color = "black", size = 1.2), dpi = 300) +
  rasterize(geom_point(size = 0.6, alpha = 0.8), dpi = 300) + 
  scale_color_manual(values = colors, 
                     name = NULL) + 
  coord_fixed(ratio = 1) + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        text = element_text(color = "black"), 
        axis.title = element_blank(), 
        axis.text = element_blank(), 
        legend.title = element_text(size = 10),
        legend.text  = element_text(size = 8)) + 
  guides(colour = guide_legend(title = "APC subclusters", override.aes = list(size = 4, alpha = 1)))

UMAP_legend <- cowplot::get_legend(UMAP)

F2D <- plot_grid(plot_grid(UMAP_PBMC, UMAP_TIL, nrow = 1, align = "hv"),
                 UMAP_legend, ncol = 2, rel_widths = c(1, 0.25))
F2D

```

## Figure 2E

UMAP of monocytes and macrophages embedded with the mean pseudotemporal value of each cell.

```{r}

data_F2E <- read.csv("../Data/Gong_et_al_Treg_Fig_2_E.csv")

x_min <- min(data_F2E$umap_1)
x_max <- max(data_F2E$umap_1)
y_min <- min(data_F2E$umap_2)
y_max <- max(data_F2E$umap_2)

F2E <- ggplot(data_F2E, aes(x = umap_1, y = umap_2, color = Avg_pseudotime_value)) + 
  rasterize(geom_point(size = 0.6), dpi = 300) + 
  scale_color_gradientn(colors = scales::alpha(viridis::magma(20)[5:20], 0.9), 
                        name = "Pseudotime", 
                        limits = c(0, 140), 
                        breaks = c(0, 35, 70, 105, 140), 
                        oob = scales::squish) + 
  coord_fixed(ratio = 1) + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        axis.title = element_blank(), 
        axis.text = element_blank()) + 
  annotate("segment", x = x_min, xend = x_min + 2, y = y_min, yend = y_min, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") +
  annotate("segment", x = x_min, xend = x_min, y = y_min, yend = y_min + 2, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") + 
  annotate("text", x = x_min + 2.5, y = y_min, 
           label = "UMAP_1", hjust = 0, vjust = 0.5, size = 4) +
  annotate("text", x = x_min, y = y_min + 2.5, 
           label = "UMAP_2", angle = 90, hjust = 0, vjust = 0.5, size = 4)
F2E

```

## Figure 2F

Spearman correlation across all tumor samples between the proportions of each Treg subcluster and the proportions of each myeloid APC subclusters.

```{r}

data_F2F <- read.csv("../Data/Gong_et_al_Treg_Fig_2_F.csv")

data_F2F$Treg_subcluster <- factor(data_F2F$Treg_subcluster, levels = Treg_levels)
data_F2F$APC_subcluster <- factor(data_F2F$APC_subcluster, levels = APC_levels)

F2F_plot_data <- data_F2F %>% 
  mutate(log_p_value = -log10(p_value)) %>% 
  mutate(significance = ifelse(p_value < 0.05, "Significant", "NS"))

F2F <- ggplot(F2F_plot_data, aes(x = Treg_subcluster, y = APC_subcluster)) +
  geom_point(aes(size = log_p_value, 
                 fill = R), 
             shape = 22, stroke = 0.2) + 
  geom_text(data = subset(F2F_plot_data, significance == "Significant"), 
            label = "*", 
            size = 5, vjust = 0.75, 
            color = "black") + 
  scale_fill_gradient2(low = "#2166AC", mid = "white",high = "#B2182B", 
                       midpoint = 0, limits = c(-1, 1), breaks = c(-1, 0, 1), 
                       labels = c("-1", "0", "1"), name = "Spearman R") +
  scale_radius(range = c(1, 17), breaks = c(0, 1.3, 2, 3, 4), 
               labels = c("1", "0.05", "0.01", "0.001", "0.0001"), 
               name = "p-value") + 
  guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5),
         size = guide_legend(title.position = "top",title.hjust = 0.5)) +
  theme_classic() + 
  theme(panel.grid = element_blank(), 
        axis.text.x = element_text(color = "black", size = 13, angle = 45, hjust = 1), 
        axis.text.y = element_text(color = "black", size = 13, angle = 0, hjust = 1), 
        legend.title = element_text(color = "black", size = 10), 
        legend.text = element_text(color = "black", size = 10), 
        legend.position = "bottom", 
        legend.box = "horizontal") +
  labs(x = NULL,
       y = NULL) + 
  coord_fixed()
F2F

```

## Figure 2G

Integrated NicheNet analysis of Treg-myeloid APC interactions.

```{r}

data_F2G <- readRDS("../Data/Gong_et_al_Treg_Fig_2_G.rds")

## Top 10 Treg-derived ligands ranked by regulatory activity


data_F2G_1_plot <- data_F2G$data_F2G_1 %>% 
  arrange(aupr_corrected) %>% 
  as.matrix()

p_ligand_aupr <- make_heatmap_ggplot(data_F2G_1_plot,
                                     "Prioritized ligands", "Ligand activity", 
                                     legend_title = "AUPR", color = "darkorange") + 
  theme(axis.text.x.top = element_blank())

## Expression frequency and levels of ligands across Treg subclusters

p_dotplot <- DotPlot(data_F2G$data_F2G_2, 
                     features = rev(rownames(data_F2G$data_F2G_1)), cols = "RdYlBu") + 
  coord_flip() +
  scale_y_discrete(position = "right")

## Quantification of ligands driving the phenotype of SPP1hi macrophages.

data_F2G_3_plot <- as.matrix(data_F2G$data_F2G_3)

p_ligand_target <- make_heatmap_ggplot(data_F2G_3_plot, 
                                       "Ligands from CD4+ Treg", 
                                       "Predicted target DEGs in SPP1-hi Macro signature",
                                       color = "purple", 
                                       legend_title = "Regulatory potential") +
  scale_fill_gradient2(low = "whitesmoke",  high = "purple")

## Integrated plot

figures_without_legend <- cowplot::plot_grid(
  p_ligand_aupr + theme(legend.position = "none"),
  p_dotplot + theme(legend.position = "none",
                    axis.ticks = element_blank(),
                    axis.title.y = element_blank(),
                    axis.title.x = element_text(size = 12),
                    axis.text.y = element_text(size = 9),
                    axis.text.x = element_text(size = 9,  angle = 90, hjust = 0)) +
    ylab("Expression in CD4+ Treg"),
  p_ligand_target + theme(legend.position = "none",
                          axis.title.y = element_blank()),
  align = "hv",
  nrow = 1,
  rel_widths = c(ncol(data_F2G_1_plot) + 6, 5 + 7, ncol(data_F2G_3_plot)))

legends <- cowplot::plot_grid(
  ggpubr::as_ggplot(ggpubr::get_legend(p_ligand_aupr)),
  ggpubr::as_ggplot(ggpubr::get_legend(p_dotplot)),
  ggpubr::as_ggplot(ggpubr::get_legend(p_ligand_target)),
  nrow = 1,
  align = "h", rel_widths = c(1.5, 1, 1))

F2G <-  cowplot::plot_grid(figures_without_legend, legends, rel_heights = c(10, 4.5), nrow = 2, align = "hv")
F2G

```

## Environment information

```{r}

sessionInfo()

```