---
title: "Gong et al Treg - Figure 4"
output: html_document
---

## Load packages

```{r}

library(tidyverse)
library(patchwork)
library(pheatmap)
library(scales)
library(ggrastr)
library(viridis)
library(cowplot)
library(ggplotify)
library(celltalker)

```

## Figure 4A

In vitro microsuppression assay (MSA) comparing the percentage of suppression on CD8+ T cells between using APCs and beads to provide TCR and co-stimulation.

```{r}

F4A <- plot_spacer()

```

## Figure 4B

Representative histograms of CD8+ T cell proliferation quantified by dilution of CellTrace Violet in three Treg/CD8 ratios.

```{r}

F4B <- plot_spacer()

```

## Figure 4C

Proportion of non-dividing live CD8+ T cells from MSA.

```{r}

F4C <- plot_spacer()

```

## Figure 4D

A total of 30,039 CD8+ T cells were bioinformatically isolated from the MSA scRNA-seq dataset and re-clustered into 8 subclusters.

```{r}

data_F4D_E_F <- read.csv("../Data/Gong_et_al_Treg_Fig_4_D_E_F.csv", row.names = 1)

## Axes ranges

x_min <- min(data_F4D_E_F$umap_1)
x_max <- max(data_F4D_E_F$umap_1)
y_min <- min(data_F4D_E_F$umap_2)
y_max <- max(data_F4D_E_F$umap_2)

## UMAP: colored by cell types

data_F4D_E_F$cell_types_fine <- as.factor(data_F4D_E_F$cell_types_fine)

colors <- setNames(hue_pal()(length(levels(data_F4D_E_F$cell_types_fine))), 
                   levels(data_F4D_E_F$cell_types_fine))

F4D <- ggplot(data_F4D_E_F, aes(x = umap_1, y = umap_2, color = cell_types_fine)) + 
  rasterize(geom_point(color = "black", size = 1.2), dpi = 300) +
  rasterize(geom_point(size = 0.6, alpha = 0.8), dpi = 300) + 
  scale_color_manual(values = colors, 
                     name = NULL) + 
  coord_fixed(ratio = 1) + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        text = element_text(color = "black"), 
        axis.title = element_blank(), 
        axis.text = element_blank(), 
        legend.title = element_text(size = 10),
        legend.text  = element_text(size = 8)) + 
  annotate("segment", x = x_min - 1, xend = x_min + 0.5, y = y_min, yend = y_min, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") +
  annotate("segment", x = x_min - 1, xend = x_min - 1, y = y_min, yend = y_min + 1.5, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") + 
  annotate("text", x = x_min + 1, y = y_min, 
           label = "UMAP_1", hjust = 0, vjust = 0.5, size = 4) +
  annotate("text", x = x_min - 1, y = y_min + 2, 
           label = "UMAP_2", angle = 90, hjust = 0, vjust = 0.5, size = 4) + 
  guides(colour = guide_legend(title = "CD8 subclusters", override.aes = list(size = 4, alpha = 1)))
F4D

```

## Figure 4E

Heatmap comparing the relative enrichment of each CD8 subcluster across four suppression conditions.

```{r}

df_F4E <- data_F4D_E_F %>% 
  select(cluster = cell_types_fine, Condition)                   # extract data
tab_F4E <- table(df_F4E$cluster, df_F4E$Condition)               # subcluster × condition contingency table
tab_prop_F4E <- prop.table(tab_F4E, margin = 1)                  # for each subcluster: calculate proportion in each condition
tab_scaled_F4E <- t(scale(t(tab_prop_F4E)))                      # row-wise (per subcluster) z-score scaling
tab_scaled_F4E[is.na(tab_scaled_F4E)] <- 0                       # remove NA
F4E <- pheatmap(mat = tab_scaled_F4E,                            # heatmap
                color = colorRampPalette(c("purple", "black", "yellow"))(1024), 
                cluster_rows = T, cluster_cols = T, 
                treeheight_row = 50, treeheight_col = 50, 
                border_color = NA, fontsize_row = 10, fontsize_col = 10)
F4E

```

## Figure 4F

Heatmap visualizing the association between each CD8 subcluster and functional states of CD8+ T cells.

```{r}

CD8_signatures <- c(
  "HALLMARK_INTERFERON_ALPHA_RESPONSE", "HALLMARK_INTERFERON_GAMMA_RESPONSE", 
  "REACTOME_COSTIMULATION_BY_THE_CD28_FAMILY", "REACTOME_PD_1_SIGNALING", "REACTOME_TCR_SIGNALING", 
  "BELTRA_texprog1", "BELTRA_texprog2", "BELTRA_texint", "BELTRA_texterm", "KUMAR_lung_cd8_up", "KUMAR_spleen_cd8_up", 
  "novershtern_naive_cd8", "novershtern_central_mem_cd8", "novershtern_eff_mem_cd8", "novershtern_eff_cd8", 
  "KEGG_CELL_CYCLE", "HALLMARK_E2F_TARGETS", "HALLMARK_G2M_CHECKPOINT")

df_F4F <- data_F4D_E_F %>%
  select(cluster = cell_types_fine, all_of(CD8_signatures))
df_F4F_mean <- df_F4F %>%
  group_by(cluster) %>%
  summarise(across(CD8_signatures, mean, na.rm = T))
mat_F4F <- df_F4F_mean %>%
  column_to_rownames("cluster") %>%
  as.matrix()
mat_scaled_F4F <- t(scale(mat_F4F))
mat_scaled_F4F[is.na(mat_scaled_F4F)] <- 0
F4F <- pheatmap(mat = mat_scaled_F4F, 
                color = colorRampPalette(c("purple", "black", "yellow"))(1024), 
                cluster_rows = T, cluster_cols = T, 
                treeheight_row = 50, treeheight_col = 50, 
                border_color = NA, fontsize_row = 10, fontsize_col = 10)
F4F

```

## Figure 4G

A total of 17,396 myeloid APCs were bioinformatically isolated from the MSA scRNA-seq dataset and re-clustered into 6 subclusters.

```{r}

data_F4G_H_I_J <- read.csv("../Data/Gong_et_al_Treg_Fig_4_G_H_I_J.csv", row.names = 1, check.names = F)

## Axes ranges

x_min <- min(data_F4G_H_I_J$umap_1)
x_max <- max(data_F4G_H_I_J$umap_1)
y_min <- min(data_F4G_H_I_J$umap_2)
y_max <- max(data_F4G_H_I_J$umap_2)

## UMAP: colored by cell types

data_F4G_H_I_J$cell_types_fine <- as.factor(data_F4G_H_I_J$cell_types_fine)

colors <- setNames(hue_pal()(length(levels(data_F4G_H_I_J$cell_types_fine))), 
                   levels(data_F4G_H_I_J$cell_types_fine))

F4G <- ggplot(data_F4G_H_I_J, aes(x = umap_1, y = umap_2, color = cell_types_fine)) + 
  rasterize(geom_point(color = "black", size = 1.2), dpi = 300) +
  rasterize(geom_point(size = 0.6, alpha = 0.8), dpi = 300) + 
  scale_color_manual(values = colors, 
                     name = NULL) + 
  coord_fixed(ratio = 1) + 
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        text = element_text(color = "black"), 
        axis.title = element_blank(), 
        axis.text = element_blank(), 
        legend.title = element_text(size = 10),
        legend.text  = element_text(size = 8)) + 
  annotate("segment", x = x_min, xend = x_min + 1.5, y = y_min, yend = y_min, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") +
  annotate("segment", x = x_min, xend = x_min, y = y_min, yend = y_min + 1.5, 
           arrow = arrow(type = "closed", angle = 20, length = unit(2, "mm"), ends = "last"), 
           linewidth = 0.6, color = "black") + 
  annotate("text", x = x_min + 2, y = y_min, 
           label = "UMAP_1", hjust = 0, vjust = 0.5, size = 4) +
  annotate("text", x = x_min, y = y_min + 2, 
           label = "UMAP_2", angle = 90, hjust = 0, vjust = 0.5, size = 4) + 
  guides(colour = guide_legend(title = "APC subclusters", override.aes = list(size = 4, alpha = 1)))
F4G

```

## Figure 4H

Heatmap comparing the relative enrichment of each myeloid subcluster across four suppression conditions.

```{r}

df_F4H <- data_F4G_H_I_J %>% 
  select(cluster = cell_types_fine, Condition)                   # extract data
tab_F4H <- table(df_F4H$cluster, df_F4H$Condition)               # subcluster × condition contingency table
tab_prop_F4H <- prop.table(tab_F4H, margin = 1)                  # for each subcluster: calculate proportion in each condition
tab_scaled_F4H <- t(scale(t(tab_prop_F4H)))                      # row-wise (per subcluster) z-score scaling
tab_scaled_F4H[is.na(tab_scaled_F4H)] <- 0                       # remove NA
F4H <- pheatmap(mat = tab_scaled_F4H,                            # heatmap
                color = colorRampPalette(c("purple", "black", "yellow"))(1024), 
                cluster_rows = T, cluster_cols = T, 
                treeheight_row = 50, treeheight_col = 50, 
                border_color = NA, fontsize_row = 10, fontsize_col = 10)
F4H

```

## Figure 4I

Heatmap visualizing the association between each myeloid subcluster and gene modules of tumor-infiltrated myeloid cells.

```{r}

TAM_signatures <- c("Classical Mono", "Non-classical Mono", 
                    "C1QC-hi Macro", "HSP-hi Macro", "Hypoxic Macro", "ISG-hi Macro", "NLRP3-hi Macro", "SPP1-hi Macro")

df_F4I <- data_F4G_H_I_J %>%
  select(cluster = cell_types_fine, all_of(TAM_signatures))
df_F4I_mean <- df_F4I %>%
  group_by(cluster) %>%
  summarise(across(TAM_signatures, mean, na.rm = T))
mat_F4I <- df_F4I_mean %>%
  column_to_rownames("cluster") %>%
  as.matrix()
mat_scaled_F4I <- t(scale(mat_F4I))
mat_scaled_F4I[is.na(mat_scaled_F4I)] <- 0
F4I <- pheatmap(mat = mat_scaled_F4I, 
                color = colorRampPalette(c("purple", "black", "yellow"))(1024), 
                cluster_rows = T, cluster_cols = T, 
                treeheight_row = 50, treeheight_col = 50, 
                border_color = NA, fontsize_row = 10, fontsize_col = 10)
F4I

```

## Figure 4J

Violin plot of ordered myeloid APC subclusters based on the mean pseudotime value of each subcluster.

```{r}

F4J <- data_F4G_H_I_J %>%
  mutate(cell_types_fine = as.factor(cell_types_fine)) %>% 
  group_by(cell_types_fine) %>% 
  mutate(`Mean pseudotime` = mean(Avg_pseudotime_value)) %>%
  ungroup() %>% 
  ggplot(.,aes(x = reorder(cell_types_fine, `Mean pseudotime`), y = Avg_pseudotime_value,
               fill = `Mean pseudotime`)) +
  geom_violin() + 
  geom_boxplot(outlier.shape = NA, width = 0.2, colour = "white") + 
  theme_bw() + 
  theme(legend.position = "bottom") + 
  xlab("APC subclusters") + 
  ylab("Pseudotime") + 
  ggtitle("CD14+CD16+ myeloid: Pseudotime value")
F4J

```

## Figure 4K

Circos plot showing top R-L pairs identified in MSA scRNA-seq dataset under the max suppression condition.

```{r}

data_F4K <- read.csv("../Data/Gong_et_al_Treg_Fig_4_K.csv")

colors_use <- hue_pal()(15)

F4K <- as.ggplot(
  ~circos_plot(ligand_receptor_frame = data_F4K,
               cell_group_colors = colors_use,
               ligand_color = "red",
               receptor_color = "purple",
               facing_inner = "clockwise", 
               facing_outer = "inside", 
               cex_inner = 0.5,
               cex_outer = 1
               )$plot
) + ggtitle("Max suppression") + 
  theme(plot.title = element_text(hjust = 0))
F4K

```

## Environment information

```{r}

sessionInfo()

```